---
title             : "Remarkable features: Using descriptive contrast to convey and infer typicality"
shorttitle        : "Descriptive contrast and typicality"

author: 
  - name          : "Claire Augusta Bergey"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "5848 S. University Avenue, Chicago, IL 60637"
    email         : "cbergey@uchicago.edu"

affiliation:
  - id            : "1"
    institution   : "The University of Chicago"


wordcount         : ""
references        : ""

bibliography      : ["dissertation.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"    
output            : papaja::apa6_pdf
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = '!tb', echo = FALSE, cache = TRUE, 
                      warning = FALSE, message = FALSE, 
                      sanitize = TRUE, fig.path='figs/', fig.width = 6,
                      fig.height = 3)
set.seed(42)
options(digits=3, dplyr.summarise.inform = FALSE)
```

```{r libraries, cache = FALSE}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(lme4)
library(broom)
library(broom.mixed)
#library(ggpubr)
library(here)
library(nnet)
library(english)
library(weights)
library(scales)
library(ggthemes)
library(papaja)
library(gridExtra)
library(glue)
library(directlabels)
library(tidyboot)
library(lmerTest)
library(knitr)
library(rwebppl)
library(ggridges)
logit <- function(x) {log(x/(1-x))}
```

```{r set-theme, cache = FALSE}
theme_set(theme_few(base_size = 10) + theme(legend.position = "none"))
```

```{r make-text-vars}
make_text_vars <- function(df, term_name, term_filter = NULL) {
  if(!is.null(term_filter)) {
    filtered_df <- df %>%
      filter(term == term_filter) 
  } else{
    filtered_df <- df
  }
    
  walk(c("estimate", "statistic", "p.value"), 
      ~assign(glue("{term_name}_{.x}"), 
              filtered_df %>% pull(!!.x), 
         envir = globalenv()))
}
```

An utterance can say much more about the world than its literal interpretation might suggest. For instance, if you hear a colleague say "We should hire a female professor," you might infer something about the speaker's goals, the makeup of a department, or even the biases of a field—none of which is literally stated. These inferences depend on recognition that a speaker's intended meaning can differ from the literal meaning of their utterance, and the process of deriving this intended meaning is called *pragmatics*. Frameworks for understanding pragmatic inference posit that speakers tend to follow general principles of conversation—for instance, that they tend to be relevant, brief, and otherwise helpfully informative [@grice1975logic; @sperber1986relevance; @clark_pragmatics_1990]. When a speaker deviates from these principles, a listener can reason about the alternative utterances the speaker might have said and infer some intended meaning that goes beyond the literal meaning of their utterance.

Beyond enriching the interpretation of utterances whose literal meaning is known, pragmatic inference is a potentially powerful mechanism for learning about new words and concepts. People can learn the meanings of words by tracking associations between word use and present objects alone [@yu2007], but reasoning about a speaker's intended meaning---not just relating the words they say to objects in the environment---may support more rapid and more accurate learning [@frank2009]. For example, @akhtar_role_1996 showed that young children can infer the meaning of a new word by using the principle that people tend to remark on things that are new and interesting to them. In this study, an experimenter leaves the room and a new toy emerges in her absence; once she comes back, the toy is familiar to the child but not to the experimenter. When she uses a novel name, "gazzer," the child can infer that the word refers to the toy that is novel to the experimenter, and not to other toys the experimenter had already seen. Experiments with adults show that they too can use general principles of informativeness to infer a novel referent's name [@frank2014].

One potential pragmatic tool for learning about referents is contrastive inference from description. To the extent that communicators strive to be minimal and informative, description should discriminate between the referent and some relevant contrasting set. This contrastive inference is fairly obvious from some types of description, such as some postnominal modifiers: "The door with the lock" clearly implies a contrasting door without one [@nietal]. The degree of contrast implied by more common descriptive forms, such as prenominal adjectives in English, is less clear: speakers do not always use prenominal adjectives minimally, often describing more than is needed to establish reference [@engelhardt_over-specified_2011; @mangold_informativeness_1988; @pechmann_incremental_1989]. Nevertheless, @sedivy_achieving_1999 showed that people can use these inferences to resolve referential ambiguity in familiar contexts. When asked to "Pick up the tall cup," people directed their attention more quickly to the target when a short cup was present, and did so in the period before they heard the word "cup." Because the speaker would not have needed to specify "tall" unless it was informative, listeners were able to use the adjective to direct their attention to a tall object with a shorter counterpart. Subsequent work using similar tasks has corroborated that people can use contrastive inferences to direct their attention among familiar referents [@sedivy_pragmatic_2003; @aparicio2016processing; @ryskin2019information].

But what if you didn't know the meaning of the key words in someone's utterance---could you use the same kind of contrastive inferences to learn about new words and categories? Suppose a friend asks you to "Pass the tall dax." Intuitively, your friend must have said the word "tall" for a reason. One possibility is that your friend wants to distinguish the dax they want from another dax they do not. In this case, you might look around the room for two similar things that vary in height, and hand the taller one to them. If, alternatively, you only see one object around whose name you don't know, you might draw a different inference: this dax might be a particularly tall dax. In this case, you might think your friend used the word "tall" for a different reason--not to distinguish the dax they want from other daxes around you, but to distinguish the dax they want from other daxes in the world. This would be consistent with data from production studies, in which people tend to describe atypical features more than they describe typical ones [@mitchell_2013; @westerbeek_2015; @rubio-fernandez_how_2016]. For instance, people almost always say "blue banana" to refer to a blue banana, but almost never say "yellow banana" to refer to a yellow one. In each of these cases---when distinguishing the dax from other referents nearby, or from daxes in general---you would have used a pragmatic inference to learn something new about the category of daxes.

This dissertation will explore the ways in which people can learn about new words and categories from contrastive inference, with an eye toward understanding how contrastive inference could help children learn about language and the world it describes. To set the stage for understanding how listeners use contrastive inference, we first need to establish that speakers use adjectives in informative ways. 

In Chapter 1, we investigate whether people tend to use adjectives to remark on the atypical features (e.g., "the purple carrot") rather than the typical features (e.g., "the [orange] carrot") of things. In a corpus study of caregivers' speech, we show that caregivers tend to mention atypical rather than typical features of things when speaking to their children. We also preliminarily ask whether adults speaking to other adults tend to remark on atypical features rather than typical ones in a large naturalistic corpus, extending findings from reference game tasks in the lab [@mitchell_2013; @westerbeek_2015; @rubio-fernandez_how_2016], and propose to investigate this question in more depth. In an analysis using language models, we also examine whether it is possible to learn about the typical features of things without pragmatic inference, using the statistical patterns within language alone. To do this, we examine whether two language models that use word co-occurrence to represent word meaning, word2vec and BERT, represent nouns as more similar to their typical adjectives than their atypical adjectives. We find that they do not: likely because they use associative methods to represent word meaning while their input tends to highlight atypical features, these models represent the relationship between nouns and adjectives poorly. We discuss implications for children's word learning as well as for language modeling. 

In Chapter 2, we will establish that adults can use contrastive inferences to learn about a new category's feature distribution. People use adjectives for multiple communicative purposes: in some cases, an adjective is needed to pick out one object among others in the immediate environment (e.g., "the tall cup" contrasts with a nearby shorter cup, but is not especially tall); in others, it marks atypicality (e.g., "the tall cup" is taller than most cups in general). In this chapter, we use two experiments with adults to show that people can use contrastive inferences to learn about a new category's feature distribution. People observe instances of novel categories and hear them described (e.g., "Pass me the [green] toma"), and then judge the prevalence of the relevant feature (e.g., how common it is for tomas to be green). People infer that mentioned features are less prevalent than unmentioned ones, and do so even when the feature had to be mentioned to establish reference. We use a model in the Rational Speech Act (RSA) framework to capture people's judgments, finding that their judgments reflect graded consideration of both reference and conveying typicality as purposes of using an adjective.

In Chapter 3, we propose to test whether children are able to use contrastive inferences to learn about the feature distributions of new categories. To do this, we will examine children's contrastive inferences about two domains: novel objects and novel social groups. We will test whether children make the inference that, for example, mentioning that a certain group member is smart, kind or strong implies that other group members are less likely to have those traits. We will also test whether mentioning that an object is round, tall or furry prompts children to make a similar inference. Investigating both of these domains allows us to understand how children come to make adult-like contrastive inferences and whether these inferences differ across categories about which children might receive more or less direct evidence. We discuss the implications of this kind of inference for children's learning given the descriptions they hear from caregivers, and the potential unintended consequences of remarking on individuals' traits for children's learning about social groups. 


# Chapter 1: People talk more about atypical than typical features of things

```{r child = "diss_sections/chapter_1/ch1.Rmd"}
```

# Chapter 2: How adults use contrastive inference to learn about new categories

```{r child = "diss_sections/chapter_2/experiment2.Rmd"}
```

```{r child = "diss_sections/chapter_2/experiment3.Rmd"}
```

## Conclusion

In Chapter 1, we established that people tend to mention atypical rather than typical features. In this chapter, we showed that adults make appropriate pragmatic inferences given how speakers describe: they infer that a mentioned feature is likely to be less typical of the mentioned category. However, the ability to learn about new categories using contrastive inference most obviously serves budding language learners---children. To fully appreciate the potential of these inferences to allow people to learn about the world, we must study their development, which we will turn to in Chapter 3.

# Chapter 3: How children use contrastive inference to learn about new categories

```{r child = "diss_sections/chapter_3/ch3.Rmd"}
```

# Conclusion

This proposal maps out a course of research to understand how speakers selectively describe remarkable features and how listeners use this selective description to learn more about the world. In doing so, it inverts the framework that has positioned pragmatic inference as augmenting literal meaning that is already known, instead considering how people can use pragmatics to learn more about the semantics of unfamiliar things. 

To understand how people use description to learn about the world, we first must know how description is used. Chapter 1 illustrates how caregivers use description in speaking to children, and proposes to more thoroughly test this question among adults speaking to adults. We find that parents predominantly mention atypical rather than typical features when speaking to children, and find the same pattern in our preliminary analysis of adult-adult speech. This pattern of description is consistent with the idea that people use language informatively with relation to background knowledge of the world, rather than giving veridical running commentary on the world's features. This finding raises questions about how children use description to learn, given that so many accounts of language learning rest on children forming associations among co-occurring words, features, and concepts. We also investigated whether language models that use associative learning among words can extract typical feature information from language. We find that these models do poorly in distinguishing between the typical and atypical features of nouns, with implications both for associative accounts of children's language learning and for language modeling. Overall, our findings highlight the complexity of learning about the world from language that describes it selectively.

However, perhaps people---unlike language models---know that language is used to selectively remark on the world, and can use this fact to learn about the unfamiliar. In Chapter 2, we investigated how adults make inferences about novel object categories, and found that they can use description to infer that a described feature is atypical. Further, even when description may have been used for another purpose---to establish reference---people still make inferences about typicality. We find that a model that considers the utility of utterances with respect to reference and typicality captures people's inferences. Much prior work has only considered the use of description in distinguishing between present referents [@pechmann_incremental_1989;  @engelhardt_over-specified_2011; @mangold_informativeness_1988], and even work that has incorporated typicality has focused on reference as the primary goal of description [@sedivy_pragmatic_2003; @mitchell_2013; @westerbeek_2015; @rubio-fernandez_how_2016]. Our findings emphasize that conveying typicality is likely a central factor in referring, and inferences about typicality are not secondary to or blocked by the purpose of establishing reference. 

The ability to exploit description to learn more about the world than one has observed directly is most useful to people who are still rapidly learning---children. In Chapter 3, we propose to investigate how children make contrastive inferences about typicality, both of novel objects and novel social groups. The results of this experiment can tell us what children learn from the description they hear, as characterized in Chapter 1. Further, examining both novel objects and social groups can tell us how these inferences affect beliefs with social consequences, and whether children weigh direct observation against information from language differently in different domains.

The core computation in pragmatic inference is reasoning about alternatives---things the speaker could have said and did not. Given that others are reasoning about these alternatives, no choice is neutral. In the studies in Chapter 2, for instance, using an adjective in referring to an object led people to infer that the feature described by that adjective was less typical than if it had not been mentioned. But, conversely, *not* using an adjective led them to think that the feature was more typical than if they could not understand the meaning of the utterance at all---all communicative choices leak one's beliefs about the world. This has implications not only for learning about novel concrete objects, as people did here, but for learning about less directly accessible entities such as abstract concepts and social groups, as we propose to examine in Chapter 3. These inferences can be framed positively, as ways for learners to extract additional knowledge that was not directly conveyed, but can also spread beliefs that the speaker does not intend. The principle that people speak informatively is simple, but it holds unintuitive consequences---among speakers and listeners, humans and machines, adults and children---for describing and learning about the world.

# Acknowledgements

Each chapter in this proposal represents collaborative work. The collaborators on each chapter are as follows: Chapter 1, Benjamin Morris and Dan Yurovsky; Chapter 2, Dan Yurovsky; Chapter 3, Rachel King and Dan Yurovsky. Thank you to my committee, Susan Goldin-Meadow, Marisa Casillas, Howard Nusbaum, and Dan Yurovsky for guidance throughout the preparation of this work.

Thank you to Ming Xiang, Benjamin Morris, Ashley Leung, Michael C. Frank, Judith Degen, Stephan Meylan, and Ruthe Foushee for feedback on portions of this research and manuscript. Portions of this work were published in the proceedings of Experiments in Linguistic Meaning (2020) and the proceedings of the 42nd annual meeting of the Cognitive Science Society. I am grateful for feedback from reviewers and attendees of Experiments in Linguistic Meaning, the meeting of the Cognitive Science Society, the meeting of the Society for Research in Child Development, the Midwestern Cognitive Science Conference, and the Dubrovnik Conference on Cognitive Science.

This research was funded by James S. McDonnell Foundation Scholar Award in Understanding Human Cognition \#220020506 to Dan Yurovsky. The funding body had no involvement in the conceptualization, data collection, or analysis of this project.


\newpage

# References

<div id = "refs"></div>
